MLISA: Machine Learning-based \\ Intelligent Service Agents

### 2014-10-31 papers/webservice-coalition-learning/webservice_coalition_learning.tex
merged babak work

\subsection{The Join Challenge}\label{s:tjc}
Sometimes web services can increase their overall utility by starting a collaboration with other web services as communities. This collaboration can help them with better ways of sharing resources, higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters and long-term outcome is dependent on these parameters.

The goal of all parties involved in the community is to maximize their long-term outcome of their collaboration. Web services need to be utilized with a selection strategy to choose from all the different possible collaboration groups they can form and also an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment different possible collaborative groups with other web services in order to be able to estimate their gain over time. However with high number of possible communities and many other business restrictions its not possible to test run collaboration with random web services. Even if a linear approximate function for estimating utility base on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of possible number of partition of web services into communities, would make any brute-force type algorithm for best community selection strategy intractable and non-practical in real-world application settings.

\subsection{Join Consequences}\label{s:jc}

@@

In this section we describe our training process and methodology assisting services with a decision making mechanisms. We first present the training process and then we evaluate the learning model of web service in community settings.

@@

(( I think this subsection is not needed? ))

@@

We used a data set that is extracted from running web services. In this data set, each web service is associated with a number of features that reflect its functionality. To engage in communities, we add additional features that reflect web service’s utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for community to be formed. Since we have 64 different time slots of extracted features for each web service, we could build synthetic data set that contains features of large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for testing web services. We have not implemented a learning model yet, but we do have all training data as well as validation and testing. So we are able to develop a classification algorithm that effectively classify web services and propose appropriate communities that best fit to web service’s requirements.

@@

Now we want to generate a set of feature vectors representing the communities of web services, for our training purposes. For communities we add two additional external features as discussed in section \ref{s:ef} that reflect the community's estimation of its quality metrics compared to other web services. Therefore we formulate a \emph{Community Feature Vector (CFV)} as $CFV_{<C>} = \{f_1,...f_5\}$ having community of $k$ web services. $C = {ws_1,...ws_k}$. The features $f_1$ through $f_5$ represent the \emph{execution time}, \emph{throughput}, \emph{availability}, the \emph{external parameter 1 and 2} correspondingly. A set of communities, with their feature vectors and utilities evaluated provides our algorithm with raw training data set. We call this set of communities the \emph{template vector} and the set of feature vectors associated with the \emph{template vector} would be the \emph{community feature vector set (CFVS)}.

\subsubsection{feature engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_n\}$ be the community feature vector set with $N$ communities. Based on $CFVS$ set we create a $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility. Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of $gain$ matrix is can be a costly process with very big number of $N$, therefore size of feature vector $|N|$ should be chosen carefully.

Now we let our set of communities in $CFVS$ set, within $|T|$ time frame iterations, choose the the best communities to join. Each community is provided with the corresponding row of data from the $gain$ matrix. Basicly $C_i$ is provided with the data in row $i$ of matrix $gain_{n,n}$ which is all the possible utility values it can gain by joining different communities. By ordering the list, each community is utilized with an ordered set of preferences over other communities it likes to join. We define $>_{i}^t$ as preference order of community $i$ at time $t$.

@@

>>>>

\subsection{The Join Challenge}\label{s:tjc}
Sometimes web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources, higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters and long-term outcome is dependent on these parameters.

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be utilized with a selection strategy to choose from all the different possible collaboration groups they can form and also an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment different possible collaborative groups with other web services in order to estimate their gain over time. However, with high number of possible communities and many other business restrictions its not possible to test run collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of possible number of partition of web services into communities, would make any brute-force type algorithm for best community selection strategy intractable and non-practical in real-world application settings.

\subsection{Join Consequences}\label{s:jc}
It is worth mentioning that a join event takes place as a result of two parties that are looking for expanding their collaborations. All actions are picked in attempt to enhance the overall outcome. However, the selected action may result in decreasing the overall utility in long term. This is the case where a single web service joins a community, but the complex process of task allocation eliminates the visibility of the web service and thus, the web service stays idle within the community. In other words the join action is not beneficial for the web service and in the first spot, it was a bad decision.  The same event might be beneficial for the community as it hosts a web service that in case of a relative task can engage the web service in the task. But overall, both sides do not benefit from collaborating with each other and the join event has negative consequences in at least one sides' utility.

There is another scenario, which is more common, that is the web service joins a community and both parties benefit from the join. This action is rational as both web service and the community web service enhance the utility. However, the community may not be the best possible choices for the web service for join. In other words, the web service could have joined a better community, but because of lack of knowledge about the surrounding environment, joins the available community that in fact enhances its utility. So the web service stays within the community for long term. We recognize these challenges and the proposed model (in the following section) provide solutions that effectively address the aforementioned challenges.

@@

In this section, we describe our training process and methodology that we quip web services with a decision making mechanism. In this methodology, we first present the training process and then we evaluate the learning model of web service in community settings.

@@

@@

We used a data set that is extracted from running web services. In this data set, each web service is associated with a number of features that reflect its functionality. These web services are opertaing in an online environment and are continuously assigned with tasks to handle. Sometime, they pass tasks around distribute sub-tasks among each other and handle the task in the form of collaborative work. To engage in communities, we add additional features that reflect web service’s utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for community to be formed. Since we have 64 different time slots of extracted features for each web service, we could build synthetic data set that contains features of large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for testing web services. To train our model, we build a decision tree as benchmark for our decision making mechanism. The decision tree is built based on training the real data we obtain from operating web services and features we extract in regards to their performance either alone as a part of joint work with other web services. As a result, we are able to classify web services to single and joined lables and propose appropriate communities that best fit to web service’s requirements.

@@

Now we want to generate a set of feature vectors representing the communities of web services, for our training purposes. For communities we add two additional external features as discussed in section \ref{s:ef} that reflect the community's estimation of its quality metrics compared to other web services. Therefore we formulate a \emph{Community Feature Vector (CFV)} as $CFV_{<C>} = \{f_1,...f_5\}$ having community of $k$ web services. $C = {ws_1,...ws_k}$. The features $f_1$ through $f_5$ represent the \emph{execution time}, \emph{throughput}, \emph{availability}, the \emph{external parameters 1 and 2} correspondingly. A set of communities, with their feature vectors and utilities evaluated provides our algorithm with raw training data set. We call this set of communities the \emph{template vector} and the set of feature vectors associated with the \emph{template vector} would be the \emph{community feature vector set (CFVS)}.

\subsubsection{feature engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on $CFVS$ set we create a $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of $gain$ matrix is can be a costly process with very big number of $N$, therefore size of feature vector $|N|$ should be chosen carefully.

Now we let our set of communities in $CFVS$ set, within $|T|$ time frame iterations, choose the the best communities to join. Each community is provided with the corresponding row of data from the $gain$ matrix. Basicly $C_i$ is provided with the data in row $i$ of matrix $gain_{N,N}$ which is all the possible utility values it can gain by joining different communities. By ordering the list, each community is utilized with an ordered set of preferences over other communities it likes to join. We define $>_{i}^t$ as preference order of community $i$ at time $t$.

@@

Having these sets built based on training model, we are able to build a decision tree where each community's preferences are set as well as the joined events that took place in different time slots. In addition to best choices, web services have access to ordered other choices and can look for the second best or third best if their first try is rejected by the target community. The consequences of each try level is analyzed in more details in the following section where we launch experiments and investigate how effective is the use of decision tree with different decision layers in joining other communities and enhancing the overall utility.
